Once we have the AST we can transform it
using the \I{Treeregexp} language.
The code below (in file \verb|I2PIR.trg|)
shows a set of algebraic tree transformations
whose goal is to produce 
machine independent optimizations.
\begin{verbatim}
{ #  Example of support code
  use List::Util qw(reduce);
  my %Op = (PLUS=>'+', MINUS => '-',
            TIMES=>'*', DIV => '/');
}
algebra = fold wxz zxw neg;

fold: /TIMES|PLUS|DIV|MINUS/:b(NUM, NUM)
=> {
  my $op = $Op{ref($b)};
  $NUM[0]->{attr} = eval
  "$NUM[0]->{attr} $op $NUM[1]->{attr}";
  $_[0] = $NUM[0];
}
zxw: TIMES(NUM, .) and {$NUM->{attr}==0}
=> { $_[0] = $NUM }
wxz: TIMES(., NUM) and {$NUM->{attr}==0}
=> { $_[0] = $NUM }
neg: NEG(NUM)
=> { $NUM->{attr} = -$NUM->{attr};
     $_[0] = $NUM }
\end{verbatim}
A Treeregexp programs is made of \I{treeregexp} rules
that describe what subtrees match and how transform them:
\begin{verbatim}
wxz: TIMES(., NUM) and {$NUM->{attr}==0}
=> { $_[0] = $NUM }
\end{verbatim}
A rule has a \I{name} (\verb|wxz| in the example), 
a \I{term} describing
the shape of the subtree to match \verb|"TIMES(., NUM)"|
and two optional fields:
a \I{semantic condition} expliciting
the attribute constraints (the code after the reserved word
\verb|and|)
and some \I{transformation code} that tells how to 
modify the subtree (the code after the big arrow \verb|=>|).
Each rule is translated into a subroutine 
\footnote{The sub must be accessed 
through a proxy 
\code{Parse::Eyapp::YATW} object. YATW stands for \I{Yet Another Tree Walker}}
with name the treerexexp rule \I{name}.
Therefore, after compilation 
a subroutine \verb|wxz| will be available.
The dot in the \I{term} \verb|TIMES(., NUM)| 
matches any tree. The semantic condition
states that the \verb|attr| entry of node
\verb|NUM| must be zero.
The \I{transformation code} - that will be 
applied only if the matching succeeded -
substitutes the whole subtree by its 
right child.

References to the nodes associated with some
\verb|CLASS| appearing in the \I{term}
section can be accessed inside the semantic parts
through the lexical variable \verb|$CLASS|.
If there is more than one node the 
associated variable is \verb|@CLASS|. Variable \verb|$_[0]|
refers to the root of the subtree that matched.

Nodes inside a \I{term} can be described using linear
regular expressions like in the \verb|fold| transformation:
\begin{verbatim}
/TIMES|PLUS|DIV|MINUS/:b(NUM, NUM)
\end{verbatim}
In such cases an optional identifier 
to later refer the node that matched 
%within semantic parts 
can be specified (\verb|b| in the example).

Tree transformations can be grouped in families:

\begin{verbatim}
algebra = fold wxz zxw neg;
\end{verbatim}

Such families - and the objects they collect - are 
available inside the client program (read anew the code
of the driver in section \ref{section:phases}). Thus,
if \verb|$t| holds the AST resulting
from the parsing phase, we can call
its method \verb|s| (for substitute)
with args the \verb|@algebra| family:
\begin{verbatim}
$t->s(our @algebra);
\end{verbatim}

The \verb|s| method of 
\verb|Parse::Eyapp::Node|\footnote{All the classes in the AST
inherit from \code{Parse::Eyapp::Node}}
proceeds to apply all the transformation in the family
\verb|@algebra| to tree \verb|$t|
until none of them matches. Thus, for input
\verb|a = 2*(a+b)*(2-4/2)| the parser
produces the tree:
\begin{verbatim}
EXPS(
  ASSIGN(
    VAR[a],
    TIMES(
      TIMES(NUM[2],PLUS(VAR[a],VAR[b])),
      MINUS(NUM[2],DIV(NUM[4],NUM[2])
    )
  )
)
\end{verbatim}
which is transformed by the call 
\verb|$t->s(@algebra)| in:
\begin{verbatim}
EXPS(ASSIGN(VAR[a],NUM[0]))
\end{verbatim}




