The code below shows the body of the grammar (file 
\code{Infix.eyp}).
Eyapp syntax very much resembles the syntax of
old cherished \code{yacc} \cite{yacc}.
An Eyapp program has three parts: \I{head}, \I{body} and \I{tail}.
Each part is separated from the former by the symbol \verb|%%|.
The head section contains declarations, code support
and directives.
The grammar rules describing
the language - and the semantic actions that indicate how
evaluate the attributes associated with the symbols -
reside in the body section.
The tail section includes Perl code that gives support
to the semantic actions. Commonly 
the lexical analyzer and error diagnostic subroutines
go there.
 

\begin{verbatim}
%right  '='        # Head section
%left   '-' '+'
%left   '*' '/'
%left   NEG
%tree 

%%
line:             # Body section
  sts <%name EXPS + ';'>
;
sts:
    %name PRINT
    PRINT leftvalue
  | exp 
;
exp:
    %name NUM    NUM
  | %name VAR    VAR
  | %name ASSIGN leftvalue '=' exp
  | %name PLUS   exp '+' exp
  | %name MINUS  exp '-' exp
  | %name TIMES  exp '*' exp
  | %name DIV    exp '/' exp
  | %name NEG
          '-' exp %prec NEG
  |            '(' exp ')'
;
leftvalue : %name VAR VAR
;
%% 
...               # tail section
\end{verbatim}

\subsection{Ambiguities and Conflicts}
The former grammar is ambiguous.
For instance, an expression like \verb|exp '-' exp| followed by a
minus \verb|'-'| can be worked in more than one way. If we
have an input like \verb|NUM - NUM - NUM| the activity of a LALR(1) parser
(the family of parsers to which Eyapp belongs)
consists of a sequence of \I{shift and reduce actions}. A \I{shift action}
has as consequence the reading of the next token. A \I{reduce action}
is finding a production rule that matches and substituting 
the \I{right hand side} (rhs) of the production by the \
\I{left hand side} (lhs).  For input \verb|NUM - NUM - NUM|
the activity will be as follows (the dot is used to indicate where the next 
input token is):


\begin{verbatim}
.NUM - NUM - NUM # shift
 NUM.- NUM - NUM # reduce exp: NUM 
 exp.- NUM - NUM # shift
 exp -.NUM - NUM # shift
 exp - NUM.- NUM # reduce exp: NUM
 exp - exp.- NUM # shift/reduce conflict
\end{verbatim}
up to this point two different decisions can be taken: the next description can be
\begin{verbatim}
 exp.- NUM # reduce by exp: exp '-' exp 
\end{verbatim}
or:
\begin{verbatim}
 exp - exp -.NUM # shift '-' 
\end{verbatim}
that is called a \I{shift-reduce conflict}: the parser must decide
whether to shift \verb|NUM| or to \I{reduce} by the rule \verb|exp: exp - exp|.

That is also the reason for the precedence declarations in the 
head section. Another kind of conflicts are \I{reduce-reduce conflicts}.
They arise when more that rhs can be applied for a reduction
action.

By associating priorities with tokens
the programmer can tell Eyapp what syntax tree
to build in case of \I{conflict}.

The declarations 
\verb|%nonassoc|, \verb|%left| and \verb|%right| 
declare and associate a {\it priority} with the tokens
that follow them.  
Tokens declared in the same line have the same precedence. 
Tokens declared in lines below have more
precedence than those declared above. Thus, in the example
we are saying that \verb|'+'| and \verb|'-'| have the same precedence
but higher than \verb|'='|. The final effect of \verb|'-'|
having greater precedence than \verb|'='| is that an
expression like \verb|a=4-5| is interpreted as \verb|a=(4-5)| 
and not as \verb|(a=4)-5|.  The use of \verb|%left| applied to \verb|'-'|
indicates that - in case of ambiguity 
and a match between precedences - 
the parser must build the tree corresponding
to a left parenthesization. Thus, \verb|4-5-9| 
is interpreted as  \verb|(4-5)-9|.

The \verb|%prec| directive can be used when
a rhs is involved in a conflict and has no tokens
inside or it has but the precedence of the last token leads
to an incorrect interpretation. A rhs can be followed by 
an optional \verb|%prec token| directive
giving the production the precedence of the \verb|token|

\begin{verbatim}
exp:   '-' exp %prec NEG { -$_[1] }
\end{verbatim}
This solves  the conflict in \verb|- NUM - NUM|
between \verb|(- NUM) - NUM| and
{\tt - (NUM - NUM)}. Since \verb|NEG| has more
priority than \verb|'-'| the first interpretation 
will win.

\subsection{Building the AST}
\verb|Parse::Eyapp| facilitates the construction of 
abstract syntax trees (AST) through the \verb|%tree|
directive. 
Nodes in the AST are blessed in the production
\verb|name|. 
A rhs can be 
{\it named} using the \verb|%name IDENTIFIER| directive. 
For each \I{rhs name} a 
class/package with name \verb|IDENTIFIER| is created. 

%Eyapp differentiates between two kinds of tokens: \I{syntactic tokens}
%and \I{semantic tokens}. 
Symbolic tokens (like \verb|NUM|
\verb|PRINT| or \verb|VAR|) 
are considered by default \I{semantic tokens}. 
String literals 
(like \verb|'+'|, \verb|'/'|, etc.)
are - unless explictly 
declared using the \code{semantic token} directive - 
considered \I{syntactic tokens}.
When building the AST syntactic tokens do not yield 
new nodes.
Semantic tokens however have their own. Thus
when feed with input \verb|b=2*a| 
the generated parser
produces the following AST\footnote{The information
between brackets shows the attribute 
for {\tt TERMINAL} nodes}:
\begin{verbatim}
EXPS(
  ASSIGN(
    VAR(TERMINAL[b]),
    TIMES(
      NUM(TERMINAL[2]),
      VAR(TERMINAL[a]))
  )
)
\end{verbatim}
Nodes of the AST are hashes that can be 
\I{decorated} with new keys/attributes.
The only reserved field is \verb|children| which is a reference to the
array of children. 
Nodes named \verb|TERMINAL| are built from the
tokens provided by the lexical analyzer. 
The couple \verb|($token, $attribute)| returned by the lexical analyzer
is stored under the keys \verb|token| and \verb|attr|.
\verb|TERMINAL| nodes also have the attribute \verb|children| which is
set to an anonymous empty list.
Observe the absence of \verb|TERMINAL| nodes corresponding to 
tokens \verb|'='| and \verb|'*'|.
If we change the status of \verb|'*'| and \verb|'='| 
to \code{semantic} using the \verb|%semantic token| directive:
\begin{verbatim}
1   %semantic token '*' '='
2   %right  '='
3   ....  etc.
\end{verbatim}
we get a - concrete - syntax tree:
\begin{verbatim}
EXPS(
  ASSIGN(
    VAR(TERMINAL[b]),
    TERMINAL[=],
    TIMES(
      NUM(TERMINAL[2]),
      TERMINAL[*],
      VAR(TERMINAL[a])
    ) # TIMES
  ) # ASSIGN
)
\end{verbatim}
Let us now consider the input \verb|2*(a+1)|.
The parser yields the tree:
\begin{verbatim}
EXPS(
  TIMES(
    NUM(
     TERMINAL[2]),
     exp_14(
       PLUS(
         VAR(TERMINAL[a]),
         NUM(TERMINAL[1]))
       ) # PLUS
  ) # TIMES
)
\end{verbatim}
Two features are noticeable: the parenthesis rule \verb|exp:| \verb|'(' exp ')'|
had no name
and got automatically one: \verb|exp_14|. The \I{name of a rhs} by 
default results from concatenating the left hand side of the rule
with the ordinal number of the rule\footnote{As it appears
in the {\tt .output} file. The {\tt .output} file can be generated 
using the {\tt -v} option of {\tt eyapp}}.
The second is that node \verb|exp_14| is useless and can be suppressed. 

The \verb|%tree| directive can be accompanied of the \verb|%bypass|
clause.  A \verb|%tree bypass| produces an automatic \I{bypass} of any
node with only one child at \I{tree-construction-time}. 
A \I{bypass operation} consists in \I{returning the only child 
of the node being visited to the father of the node and re-typing (re-blessing)
the node in the name of the production}\footnote{If the production has an
explicit name. Otherwise there is no re-blessing}. 

Changing the line \verb|%tree| by \verb|%tree bypass|
in file \verb|Infix.eyp| we get a more suitable 
AST for input \verb|2*(a+1)|:
 
\begin{verbatim}
EXPS(TIMES(NUM[2],PLUS(VAR[a],NUM[1])))
\end{verbatim}

The node \verb|exp_14| has disapeared in this version
since the \I{bypass operation} applies to the rhs 
of the rule \verb|exp: '(' exp ')'|:
Tokens \verb|'('| and \verb|')'| are syntactic tokens
and therefore at \I{tree construction time}
only one child is left. Observe also the absence 
of \verb|TERMINAL| nodes. Bypass clearly applies
to rules \verb|exp: NUM| and \verb|exp: VAR| since
they have only one element on their rhs. Therefore the
\verb|TERMINAL| node is re-blessed as \verb|NUM| and
\verb|VAR| respectively.


A consequence of the global scope application of \verb|%tree bypass|
is that undesired bypasses may occur. Consider the tree 
rendered for input \verb|-a*2|:

\begin{verbatim}
EXPS(TIMES(NEG,NUM))
\end{verbatim}

What happened? The bypass is applied to the rhs 
\verb|'-' exp|.  Though the rhs has two symbols, token \verb|'-'| is
a syntactic token and at \I{tree-construction-time}
only \verb|exp| is left. The \I{bypass}
operation applies when building this node.
This undesired \I{bypass} can be avoided applying 
the \verb|no bypass| directive to the 
production:

\begin{verbatim}
 exp : %no bypass NEG
       '-' exp %prec NEG
\end{verbatim}
Now the AST for \verb|-a*2| is correct:
\begin{verbatim}
EXPS(TIMES(NEG(VAR),NUM))
\end{verbatim}

Eyapp provides operators \verb|+|, \verb|*| and \verb|?| 
for the creation of lists and optionals as in:
\begin{verbatim}
line: sts <EXPS + ';'>
\end{verbatim}
which states that a \code{line} is made of a non empty
list of \verb|EXPS| separated by semicolons.
By default the class name for such list is \verb|_PLUS_LIST|.
The \verb|%name| directive can be used to modify
the default name:
\begin{verbatim}
line: sts <%name EXPS + ';'>
\end{verbatim}

Explicit actions can be specified by the programmer. 
They are managed as anonymous subroutines
that receive as arguments the attributes of the symbols
in the rule and are executed each time a \I{reduction}
by that rule occurs. When running under the \verb|%tree| directive
this provides a mechanism to influence the shape of the AST.
Observe however that the grammar in the example is \underline{clean} of actions:
\I{Parse::Eyapp allowed us to produce a suitable AST without writing 
any explicit actions}.
