%
%  $Description: Author guidelines and sample document in LaTeX 2.09$
%
%  $Author: ienne $
%  $Date: 1995/09/15 15:20:59 $
%  $Revision: 1.4 $
%
\documentclass[times, 10pt,twocolumn]{article}
\usepackage{latex8}
\usepackage{times, graphics}
%\documentstyle[times,art10,twocolumn,latex8]{article}
%-------------------------------------------------------------------------
% take the % away on next line to produce the final camera-ready version
\pagestyle{empty}
%-------------------------------------------------------------------------
\textwidth 17 cm
\textheight 256 mm
\columnsep 3 mm

\newcommand{\code}[1]{{\tt #1}}
\newcommand{\I}[1]{{\it #1}}
%\newenvironment{tabcode}{
%\begin{center}
%\begin{tabular}{p{7cm}}
%}
%{
%\end{tabular}
%\end{center}
%}
%\newcommand{\ici}[1] {\index{#1}}

\begin{document}
\title{Parsing Strings and Trees with {\tt Parse::Eyapp}\\
\it (An Introduction to Compiler Construction)}
\author{
Casiano Rodriguez-Leon\\
Dpto. Estad\'{\i}stica, I.O. y Computaci\'{o}n\\
Universidad de La Laguna \\ La Laguna, 38271, Spain\\ casiano@ull.es\\
}
\maketitle
%
% if the list of authors exceeds the space for a headline,
% an abbreviated author list is needed

\begin{abstract}
Parse::Eyapp (Extended yapp) is a collection of modules
that extends Francois Desarmenien Parse::Yapp 1.05.
Eyapp extends yacc/yapp syntax with 
functionalities like named attributes,
EBNF-like expressions, modifiable default action,
automatic syntax tree building,
semi-automatic abstract syntax tree building,
translation schemes, tree regular expressions,
tree transformations, scope analysis support,
and directed acyclic graphs among others.
This article teaches you the basics of 
Compiler Construction and Parse::Eyapp by
building a translator from infix expressions to Parrot 
Intermediate Representation.
\end{abstract}
%-------------------------------------------------------------------------------


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{section:introduction}
\input{introduction.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Phases of a Translator}
\label{section:phases}
\input{phases.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lexical Analysis}
\label{section:lexicalanalysis}
\input{lexicalanalysis.tex}

%-------------------------------------------------------------------------------
%\label{section:errordiagnostics}
%\input{errordiagnostics.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Syntax Analysis}
\label{section:syntacticanalysis}
\input{syntacticanalysis.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tree Transformations}
\label{section:machineindependentoptimizations}
\input{machineindependentoptimizations.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Resource Allocation}
\label{section:addressassign}
\input{addressassign.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code Generation}
\label{section:codegeneration}
\input{codegeneration.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Peephole Transformations}
\label{section:peephole}
\input{peephole.tex}

\section{Output Generation}
\label{section:output}
\input{output.tex}


\section{Conclusions and Future Work}
There is a shortage of compiler toolkits in CPAN/Perl 5.
It will be beneficial to have a CPAN wider covering of 
other compiler components: code generators as iburg \cite{iburg} to mention 
just one.

This work presented \code{Parse::Eyapp}, a work in progress in that direction.
\code{Yacc} and \code{Parse::Yapp} programmers will feel at home
in \code{Parse::Eyapp}. 
Additionally to the beneficial mature approach to parsing provided by
\code{Yacc}-like parser generators,
\code{Parse::Eyapp} delivers a set of extensions
that give support to the later phases
of text processing.

\section{About the Author}
Casiano Rodriguez-Leon is a Professor of Computer Science at Universidad
de La Laguna. His research focuses on Parallel Computing. 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}

This work has been supported by the {\sc ec (feder)} and by
the Spanish Ministry of Education and Science inside the
`Plan Nacional de {\sc i+d+}i' with contract number {\sc tin2005-08818-c04-04}.
Thanks To Francois Desarmenien. \code{Parse::Eyapp} shares a large percentage 
of code with \code{Parse::Yapp}.


\bibliographystyle{unsrt}
\bibliography{perl}

\label{end:of:file}

\end{document}
